\chapter{The Interstellar Problem}
This chapter presents the design, implementation and analysis of my own original dynamic programming problems, The Interstellar Problem I and The Interstellar Problem II.
The problem reads as follows:\\

The robot TARS has been sucked into the singularity, an N-dimensional grid. He starts in a corner of the grid. In order to escape, he must reach the opposite corner of the singularity from where he started.
Inside, he can move in any direction through any dimension, as long as he is making progress towards his escape point with each move.
\begin{enumerate}
    \item How many different paths could TARS possibly take through the singularity in order to reach his goal and escape?
    \item If each space in the singularity has a fuel cost, what is the minimum cost path from the start position to the goal?
\end{enumerate}

\section{The Interstellar Problem I}
\begin{description}
    \item[Problem Statement:]
    Given as input the shape of an n-rank tensor, find the number of unique paths from a corner of the tensor denoted $R$ to the opposite corner of the tensor,
    denoted $F$, where the only moves allowed at any point in time strictly make progress towards $F$.
        
    \item[Input:]
    A tuple of integers $dimensions$, where $|dimensions|$ is the number of dimensions in the n-rank tensor, and each integer in $dimensions$ represents the size of the corresponding dimension.
        
    \item[Output:] 
    An integer $paths$, which represents the number of unique paths (where negative progress is not allowed) from $R$ to $F$.
        
    \item[Example 1:] For:\\
    $dimensions$ = $(10)$\\
    $paths = 1$
    \item[Explanation 1:]
    There is only one dimension in the singularity, hence one path:
    \begin{table}[H]
        \centering
        \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
            \hline
            \textbf{R} & $\rightarrow$ & $\rightarrow$ &$\rightarrow$ & $\rightarrow$ & $\rightarrow$& $\rightarrow$&$\rightarrow$ & \textbf{F} \\
            \hline
        \end{tabular}
        \caption{Visualization for Example 1}
    \end{table}

    \item[Example 2:] For:\\
    $dimensions$ = $(3,7)$\\
    $paths = 28$

    \item[Explanation 2:]
    In this case the singularity has two dimensions, and the shape is $(3,7)$. This means TARS can go right or down, as any other movement would be negative progress.
    There are 28 unique ways the robot can reach $F$ from $R$.
    Example Path:

    \begin{table}[htbp]
        \centering
        \begin{tabular}{|c|c|c|c|c|c|c|}
            \hline
            \textbf{R} & $\rightarrow$ & $\downarrow$ &  &  &  &  \\
            \hline
             &  & $\rightarrow$ & $\rightarrow$ & $\rightarrow$ & $\rightarrow$ & $\downarrow$ \\
            \hline
             &  &  &  &  &  & \textbf{F} \\
            \hline
        \end{tabular}
        \caption{Visualization for Example 2}
    \end{table}

    \item[Example 3:] For:\\
    $dimensions$ = $(2,7,5,2)$\\
    $paths = 27720$

    \item[Explanation 3:]
    In this case the singularity has 4 dimensions, making it impossible to visualize. There are 27720 different paths TARS could take to reach $F$ from $R$.
\end{description}

\section{The Interstellar Problem II}
\begin{description}
    \item[Problem Statement:]
    Given as input an n-rank tensor where the value at each cell represents the fuel cost of landing on that cell, find the cost of the path from a corner of the tensor denoted $R$ to the opposite corner of the tensor,
    denoted $F$ with the minimum total fuel cost, where the only moves allowed at any point in time strictly make progress towards $F$.
        
    \item[Input:]
    An n-rank tensor of integers called $cost$, where each cell in $cost$ represents the fuel cost of landing on that cell.
        
    \item[Output:] 
    An integer $min\_path$, which represents the cost of the cheapest path (where negative progress is not allowed) from $R$ to $F$.
        
    \item[Example 1:] For:\\
    $cost$ = $[10]$\\
    $min\_path = 10$
    \item[Explanation 1:]
    The singularity is a single cell of cost 10, so the total cost of the minimum path is 10:

    \item[Example 2:] For:\\
    $cost = $\\
    $[[1,3,1],$\\
    $[1,5,1],$\\
    $[4,2,1]]$\\
    $min\_path = 7$

    \item[Explanation 2:]
        Explanation: $1 + 3 + 1 + 1 + 1 = 7$ is the minimum path cost through the singularity.
        \begin{table}[H]
            \centering
            \begin{tabular}{|c|c|c|}
                \hline
                \textbf{R} & $\rightarrow$ & $\downarrow$ \\
                \hline
                 &  & $\downarrow$ \\
                \hline
                 &  & \textbf{F} \\
                \hline
            \end{tabular}
            \caption{Visualization for Example 2}
        \end{table}
\end{description}

\section{Building Up to The Interstellar Problem}
In order to understand the solution to the interstellar problem, we must first understand the following smaller problems,
and build up to the solution to the interstellar problem.

\subsection{Unique Paths}
\input{chapters/algorithms/al-unique-paths.tex}

\subsection{Min Path Sum}
\input{chapters/algorithms/al-min-path-sum.tex}

\subsection{Extending Pathfinding Problems to 3 Dimensions}
\input{chapters/algorithms/al-3d-unique-paths.tex}

\subsection{Correctness of N-Dimensional Pathfinding Problems}
\input{chapters/algorithms/al-nd-unique-paths.tex}